---
title: "GSS challange More Data"
author: "Anthony Falconio"
date: "Fall 2023"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

$H_0$ There is no significant difference in the data

$H_A$ There is a significant difference in the data

```{r eval=FALSE}
mydata <- mydata %>%
mutate(across(where(is.character), as.factor))
```

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

I am going to use the CandP helper files to help me compare two categorical types of data called cappun and race. A bar chart will be used in this case as it involves a comparison of cat-car. Fav stats will also be run after the fact to show both the graphical and numerical results.

$\Huge\chi^2=\sum\frac{(0-E)^2}{0}$

A chi-squared test will then be run under the inferential results to see if there is a significant difference in the data. Since there are more than two factors in the variable "race" a fisherXY function will be used with the fisher test to analyze the difference in the data if present.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~race + cappun,data=mydata)
barchartGC(~race + cappun,data=mydata, type="percent")
```

These two bar charts representative of the data show that the number of people in the survey skewed mostly towards the white race with a higher frequency in general while most within that race support capital punishment while the black and other race are against capital punishment.

The second bar chart shows the percentages within each race that either favor or oppose capital punishment. The white race is consistent with the first bar graph as the majority favor capital punishment while the minority oppose as is the same with the black race with the majority opposing it and the minority favoring it. The main difference comes from the other race in the second graph as it actually shows that the majority within this group are in favor of capital punishment while the minority oppose it.

## Now with The Most data (fewer variables gives more data points)



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
```

This table shows data consistent with the second bar chart show as the percentages match in terms of race and preference of favoring or opposing capital punishment.

## Now with More data



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
```
```{r}
fisherXY(mydata$race,mydata$cappun,printAll = TRUE)
```

The Chi-squared test shows that the p-value within the data is representative of a significant difference in the data between the races when it comes to favoring or opposing capital punishment. The observed counts also correlate with the counts expected by null. The fisher XY test corroborates the p-value and the data table showing the differences in races between opinions on capital punishment.

## More data version


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
